---
title: "Predictive Analytics Project Exploratory 11-17"
author: "JLewyckyj"
date: "11/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Data

```{r}
setwd("/nfs/home/jsl2985/z/Predictive Analytics/Projects")

book <- read.csv("book.csv")

train <- subset(book, !is.na(book$logtargamt))
test <- subset(book, is.na(book$logtargamt))
```

#Correlations between category variables, all data

```{r}
f <- book[,c(7:36)]
m <- book[,c(37:66)]

cor1 <- cor(f)
cor2 <- cor(m)

write.csv(cor1, "Frequency Correlations.csv")
write.csv(cor2, "Amount Correlations.csv")
```

Frequencies Correlations:<br>
Fiction is pretty correlated with many categories<br>
  Highest are Classics (0.52), Cartoon (0.51), Hobby (0.45), Art (0.44), History (0.44)<br>
  <br>
Philosophy and Religion: 0.47<br>
History and Religion: 0.49<br>
History and Art: 0.54<br>
History and Contemporary History: 0.51<br>
History and Travel Guides: 0.45<br>
Art and Travel Guides: 0.43<br>
<br>
Linguistics and Encyclopedia: 0.48
Health and Learning: 0.49<br>
Encyclopedia and Learning: 0.46<br>
<br>
Hobby and Nature: 0.53<br>
<br>
Uncorrelated with everything: Facsimile, Sports, Video, Non-Books
<br><br>
Amounts Correlations:<br>
Fiction is pretty correlated with many categories<br>
  Highest are Cartoons (0.39), Art (0.35), History (0.34)<br>
  <br>
Art and History: 0.48<br>
Art and Travel Guides: 0.39<br>
History and Contemporary History: 0.42<br>
History and Travel Guides: 0.38<br>
<br>
Hobby and Nature: 0.40
Uncorrelated with everything: Religion, Music, Facsimile, Health, Learning, Sports, Non-Books
<br><br>
History, Art, Contemporary History, Travel Guides correlated in both frequency and amounts, as are Hobby and Nature, and Fiction and Cartoon

#Relationships between category variables and logtargamt, only training set

```{r}
data_corr <- train[,-c(1,3:6)]
cor(data_corr[-1], data_corr$logtargamt)
```

Very uncorrelated for each category on logtargamt.

##Correlations only for logtargamt > 0

```{r}
train$successFlag <- 0
train$successFlag[train$logtargamt > 0] = 1

data_corr <- train[train$successFlag == 1,-c(1,3:6)]
cor(data_corr[-1], data_corr$logtargamt)
```

