---
title: "Saurabh1"
author: "JLewyckyj"
date: "11/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading the files and importing libraries

```{r}
setwd("/nfs/home/jsl2985/z/Predictive Analytics/Projects")
library(lubridate)

book <- read.csv("book.csv")
ordersall <- read.csv("ordersall.csv", stringsAsFactors = FALSE)

summary(book)
summary(ordersall)
```

# Coding success flag in the book database and dropping unnecessary variables
```{r}
book$successFlag <- 0
book$successFlag[book$logtargamt > 0] = 1

book <- book[,c('id', 'logtargamt', 'recency', 'frequency', 'amount', 'tof', 'successFlag')]
summary(book)
```
# Extracting features from the ordersall dataset
```{r}
#Converting the orderdate column to date format
typeof(ordersall$orddate)
head(ordersall)
ordersall$orddate <- dmy(ordersall$orddate)
head(ordersall)

#Creating a 3 months database
ordersall3mo <- ordersall[ordersall$orddate >= dmy('1-May-14'),]
summary(ordersall3mo)
aggordersall3mo <- ordersall3mo[,c('id', 'qty', 'price')]
aggordersall3mo <- aggregate(aggordersall3mo, by = list(aggordersall3mo$id), FUN = sum)
aggordersall3mo$id <- aggordersall3mo$Group.1
aggordersall3mo <- aggordersall3mo[,c('id', 'qty', 'price')]
summary(aggordersall3mo)

#Creating a 6 months database
ordersall6mo <- ordersall[ordersall$orddate >= dmy('1-Feb-14'),]
summary(ordersall6mo)
aggordersall6mo <- ordersall6mo[,c('id', 'qty', 'price')]
aggordersall6mo <- aggregate(aggordersall6mo, by = list(aggordersall6mo$id), FUN = sum)
aggordersall6mo$id <- aggordersall6mo$Group.1
aggordersall6mo <- aggordersall6mo[,c('id', 'qty', 'price')]
summary(aggordersall6mo)
```

#Merging and creating new features

```{r}

bookMod <- merge(book, aggordersall3mo, by = 'id', all.x = TRUE, all.y = FALSE)
bookMod <- merge(bookMod, aggordersall6mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('3mo','6mo'))
#summary(book)
#summary(bookMod)

bookMod$qty3mo[is.na(bookMod$qty3mo)] <- 0
bookMod$qty6mo[is.na(bookMod$qty6mo)] <- 0
bookMod$price3mo[is.na(bookMod$price3mo)] <- 0
bookMod$price6mo[is.na(bookMod$price6mo)] <- 0

#Creating a new user flag
bookMod$newUserFlag <- 0
bookMod$newUserFlag[bookMod$tof == 0] <- 1


#Creating a flag to track if the user has ordered in the past 1 month
bookMod$orderinrecent1mo <- 0
bookMod$orderinrecent1mo[(bookMod$recency <= 30) && (bookMod$tof > 0)] <- 1


#Creating a flag to track if the user has ordered in the past 3 months
bookMod$orderinrecent3mo <- 0
bookMod$orderinrecent3mo[(bookMod$recency <= 90) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user has ordered in the past 6 months
bookMod$orderinrecent6mo <- 0
bookMod$orderinrecent6mo[(bookMod$recency <= 180) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user started using the website in the past 30 days
bookMod$startedinrecent1mo <- 0
bookMod$startedinrecent1mo[(bookMod$tof <= 30) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user started using the website in the past 90 days
bookMod$startedinrecent3mo <- 0
bookMod$startedinrecent3mo[(bookMod$tof <= 90) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user started using the website in the past 180 days
bookMod$startedinrecent6mo <- 0
bookMod$startedinrecent6mo[(bookMod$tof <= 180) & (bookMod$tof != 0)] <- 1

#Creating a field to calculate % of total sales in the past 3 month
bookMod$PerSales3mo <- bookMod$price3mo / bookMod$amount * 100.0
bookMod$PerSales3mo[is.na(bookMod$PerSales3mo)] <- 0


#Creating a field to calculate % of total sales in the past 6 month
bookMod$PerSales6mo <- bookMod$price6mo / bookMod$amount * 100.0
bookMod$PerSales6mo[is.na(bookMod$PerSales6mo)] <- 0

summary(bookMod)

```

#Preliminary logistic regression model on the created dataset

```{r}
# Separating out the test from training dataset

bookModTrain <- bookMod[(!is.na(bookMod$logtargamt)),]

bookglm <- glm(successFlag ~ 
                 recency + 
                 frequency + 
                 amount +
                 tof +
                 qty3mo + 
                 price3mo + 
                 qty6mo +
                 price6mo +
                 orderinrecent1mo +
                 orderinrecent3mo +
                 orderinrecent6mo +
                 startedinrecent1mo +
                 startedinrecent3mo +
                 startedinrecent6mo +
                 PerSales3mo +
                 PerSales6mo,
               data = bookModTrain,
               family = binomial)


summary(bookglm)

```

#JL: Linear Regression Model version of above Logit model

```{r}
bookModTrain2 <- subset(bookModTrain, logtargamt > 0)

booklm <- lm(logtargamt ~ 
               recency + 
                 frequency + 
                 amount +
                 tof +
                 qty3mo + 
                 price3mo + 
                 qty6mo +
                 price6mo +
                 orderinrecent1mo +
                 orderinrecent3mo +
                 orderinrecent6mo +
                 startedinrecent1mo +
                 startedinrecent3mo +
                 startedinrecent6mo +
                 PerSales3mo +
                 PerSales6mo,
               data = bookModTrain2)

summary(booklm)
```

In Logit, significant variables are:<br>
recency, tof, frequency at p < 1%<br>
amount, PerSales6mo at p < 5% <br>
price6mo at p < 10%<br><br>

In Linear, significant variables are:<br>
frequency, amount at p < 1%<br>
qty3mo, orderinrecent3mo at p <10%

#Multicollinearity in Linear Model

```{r}
cor(bookModTrain2[,c(2,3,4,5,6)])
```

amount and frequency are highly correlated<br>
tof is fairly correlated with frequency, amount, and recency