---
title: "MSiA401Team1Project"
author: "Team 1"
date: "11 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading the files and importing libraries

```{r}
rm(list = ls())
library(pROC)
library(lubridate)

book <- read.csv("book.csv")
ordersall <- read.csv("ordersall.csv", stringsAsFactors = FALSE)
booktest <- read.csv("booktest.csv", stringsAsFactors = FALSE)

summary(book)
summary(ordersall)
```

# Coding success flag in the book database and dropping unnecessary variables
```{r}
book$successFlag <- 0
book$successFlag[book$logtargamt > 0] = 1

book$Mgroup <- book$Mfiction1 + 
  book$Mcartoons5 + 
  book$Mart12 + 
  book$Mhistory19 + 
  book$Mtravelguides31 + 
  book$Mhobby40 + 
  book$Mconthist20 + 
  book$Mnature41

book$Fgroup <- book$Ffiction1 + 
  book$Fcartoons5 + 
  book$Fart12 + 
  book$Fhistory19 + 
  book$Ftravelguides31 + 
  book$Fhobby40 + 
  book$Fconthist20 + 
  book$Fnature41


#book <- book[,c('id', 'logtargamt', 'recency', 'frequency', 'amount', 'tof', 'successFlag', 'Mgroup', 'Fgroup')]
#summary(book)
```
# Extracting features from the ordersall dataset
```{r}
#Converting the orderdate column to date format
typeof(ordersall$orddate)
head(ordersall)
ordersall$orddate <- dmy(ordersall$orddate)
head(ordersall)

#Creating a total quantity database
ordersalltot <- ordersall
aggordersalltot <- ordersalltot[,c('id', 'qty')]
aggordersalltot <- aggregate(aggordersalltot, by = list(aggordersalltot$id), FUN = sum)
aggordersalltot$id <- aggordersalltot$Group.1
aggordersalltot <- aggordersalltot[,c('id', 'qty')]
summary(aggordersalltot)


#Creating a 1 months database
ordersall1mo <- ordersall[ordersall$orddate >= dmy('1-Jul-14'),]
#summary(ordersall1mo)
aggordersall1mo <- ordersall1mo[,c('id', 'qty', 'price')]
aggordersall1mo <- aggregate(aggordersall1mo, by = list(aggordersall1mo$id), FUN = sum)
aggordersall1mo$id <- aggordersall1mo$Group.1
aggordersall1mo <- aggordersall1mo[,c('id', 'qty', 'price')]
summary(aggordersall1mo)
names(aggordersall1mo) <- c('id', 'qty1mo', 'price1mo')


#Creating a 3 months database
ordersall3mo <- ordersall[ordersall$orddate >= dmy('1-May-14'),]
#summary(ordersall3mo)
aggordersall3mo <- ordersall3mo[,c('id', 'qty', 'price')]
aggordersall3mo <- aggregate(aggordersall3mo, by = list(aggordersall3mo$id), FUN = sum)
aggordersall3mo$id <- aggordersall3mo$Group.1
aggordersall3mo <- aggordersall3mo[,c('id', 'qty', 'price')]
summary(aggordersall3mo)

#Creating a 6 months database
ordersall6mo <- ordersall[ordersall$orddate >= dmy('1-Feb-14'),]
#summary(ordersall6mo)
aggordersall6mo <- ordersall6mo[,c('id', 'qty', 'price')]
aggordersall6mo <- aggregate(aggordersall6mo, by = list(aggordersall6mo$id), FUN = sum)
aggordersall6mo$id <- aggordersall6mo$Group.1
aggordersall6mo <- aggordersall6mo[,c('id', 'qty', 'price')]
summary(aggordersall6mo)

#Creating a 12 months database
ordersall12mo <- ordersall[ordersall$orddate >= dmy('1-Aug-13'),]
#summary(ordersall12mo)
aggordersall12mo <- ordersall12mo[,c('id', 'qty', 'price')]
aggordersall12mo <- aggregate(aggordersall12mo, by = list(aggordersall12mo$id), FUN = sum)
aggordersall12mo$id <- aggordersall12mo$Group.1
aggordersall12mo <- aggordersall12mo[,c('id', 'qty', 'price')]
summary(aggordersall12mo)
names(aggordersall12mo) <- c('id', 'qty12mo', 'price12mo')
```

#Merging and creating new features

```{r}

bookMod <- merge(book, aggordersall3mo, by = 'id', all.x = TRUE, all.y = FALSE)
bookMod <- merge(bookMod, aggordersall6mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('3mo','6mo'))
bookMod <- merge(bookMod, aggordersall1mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('','1mo'))
bookMod <- merge(bookMod, aggordersall12mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('','12mo'))
bookMod <- merge(bookMod, aggordersalltot, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('','tot'))

#summary(book)
#summary(bookMod)

bookMod$qty3mo[is.na(bookMod$qty3mo)] <- 0
bookMod$qty6mo[is.na(bookMod$qty6mo)] <- 0
bookMod$qty1mo[is.na(bookMod$qty1mo)] <- 0
bookMod$qty12mo[is.na(bookMod$qty12mo)] <- 0
bookMod$price3mo[is.na(bookMod$price3mo)] <- 0
bookMod$price6mo[is.na(bookMod$price6mo)] <- 0
bookMod$price1mo[is.na(bookMod$price1mo)] <- 0
bookMod$price12mo[is.na(bookMod$price12mo)] <- 0
bookMod$qty[is.na(bookMod$qty)] <- 0


#Creating a new user flag
bookMod$newUserFlag <- 0
bookMod$newUserFlag[bookMod$tof == 0] <- 1

#Creating a onetothreemoflag month flag
bookMod$onetothreeflagmo <- 0  
bookMod$onetothreeflagmo[((bookMod$qty3mo - bookMod$qty1mo) > 0) & (bookMod$qty1mo == 0)] <- 1

table(bookMod$onetothreeflag)

#Creating a flag to track if the user has ordered in the past 1 month
bookMod$orderinrecent1mo <- 0
bookMod$orderinrecent1mo[(bookMod$recency <= 30) && (bookMod$tof > 0)] <- 1


#Creating a flag to track if the user has ordered in the past 3 months
bookMod$orderinrecent3mo <- 0
bookMod$orderinrecent3mo[(bookMod$recency <= 90) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user has ordered in the past 6 months
bookMod$orderinrecent6mo <- 0
bookMod$orderinrecent6mo[(bookMod$recency <= 180) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user has ordered in the past 12 months
bookMod$orderinrecent12mo <- 0
bookMod$orderinrecent12mo[(bookMod$recency <= 360) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user started using the website in the past 30 days
bookMod$startedinrecent1mo <- 0
bookMod$startedinrecent1mo[(bookMod$tof <= 30) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user started using the website in the past 90 days
bookMod$startedinrecent3mo <- 0
bookMod$startedinrecent3mo[(bookMod$tof <= 90) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user started using the website in the past 180 days
bookMod$startedinrecent6mo <- 0
bookMod$startedinrecent6mo[(bookMod$tof <= 180) & (bookMod$tof != 0)] <- 1


#Creating a flag to track if the user started using the website in the past 360 days
bookMod$startedinrecent12mo <- 0
bookMod$startedinrecent12mo[(bookMod$tof <= 360) & (bookMod$tof != 0)] <- 1


#Creating a field to calculate % of total sales in the past 3 month
bookMod$PerSales3mo <- bookMod$price3mo / bookMod$amount * 100.0
bookMod$PerSales3mo[is.na(bookMod$PerSales3mo)] <- 0


#Creating a field to calculate % of total sales in the past 6 month
bookMod$PerSales6mo <- bookMod$price6mo / bookMod$amount * 100.0
bookMod$PerSales6mo[is.na(bookMod$PerSales6mo)] <- 0

#summary(bookMod)

bookMod$amtuponorders <- bookMod$amount / bookMod$frequency
bookMod$qtyuponorders <- bookMod$qty / bookMod$frequency
bookMod$amtuponqty <- bookMod$amount / bookMod$qty
bookMod$ordersupontof <- bookMod$frequency / bookMod$tof
bookMod$amtupontof <- bookMod$amount / bookMod$tof
bookMod$qtyupontof <- bookMod$qty / bookMod$tof

bookMod$amtuponorders[is.na(bookMod$amtuponorders)] <- 0
bookMod$qtyuponorders[is.na(bookMod$qtyuponorders)] <- 0
bookMod$amtuponqty[is.na(bookMod$amtuponqty)] <- 0
bookMod$ordersupontof[is.na(bookMod$ordersupontof)] <- 0
bookMod$amtupontof[is.na(bookMod$amtupontof)] <- 0
bookMod$qtyupontof[is.na(bookMod$qtyupontof)] <- 0

dim(bookMod)

dim(bookMod)
summary(bookMod)

dim(bookMod[(is.infinite(bookMod$PerChangeQty3mo)) & (!is.na(bookMod$logtargamt)),])

```
# Removing Outliers

```{r}

bookModwo <- bookMod

bookModwo2 <- bookModwo[bookModwo$amtuponorders < 1000,]

bookModwo3 <- bookModwo2[bookModwo2$qtyuponorders < 40,]

bookModwo4 <- bookModwo3[bookModwo3$amtuponqty < 600,]


plot(bookModwo4$amtuponorders)
plot(bookModwo4$qtyuponorders)
plot(bookModwo4$amtuponqty)

dim(bookMod)
summary(bookModwo4)

perRemoval = (dim(bookMod)[1] - dim(bookModwo4)[1]) / dim(bookMod)[1] * 100

```

#Preliminary logistic regression model on the created dataset

```{r}
# Separating out the test from training dataset

#---------------------------------------------------------------------------------

#Without outlier removed select this code line
#bookModTrain <- bookMod[(!is.na(bookMod$logtargamt)),]
str(bookModTrain)



#With outliers removed select this line
bookModTrain <- bookModwo4[(!is.na(bookModwo4$logtargamt)),]

# Removing tof = 0
bookModTrain <- bookModTrain[bookModTrain$tof > 0,]

#str(bookModTrain)
#str(bookModTrain[bookModTrain$successFlag == 1,])

InitialFactor = dim(bookModTrain[bookModTrain$successFlag == 1,])[1]

#---------------------------------------------------------------------------------


#This data will be used for linear regression
bookModTrain2 <- subset(bookModTrain, logtargamt > 0)
dim(bookModTrain2)

#---------------------------------------------------------------------------------
# Bootstraping the data

w = ifelse(bookModTrain$successFlag == 0, 0.8 / length(which(bookModTrain$successFlag == 0)) * 10000, 0.2 / length(which(bookModTrain$successFlag == 1)) * 10000)
bs_idx = sample(seq_len(nrow(bookModTrain)), nrow(bookModTrain), prob=w, replace = TRUE)
bookModTrain = bookModTrain[bs_idx, ]

#str(bookModTrain)
#str(bookModTrain[bookModTrain$successFlag == 1,])

ScalingFactor <- 1.0 * (dim(bookModTrain[bookModTrain$successFlag == 1,])[1] - InitialFactor) / InitialFactor

# Running the logistic regression
bookglm1 <- glm(successFlag ~ 
                  recency +
                  frequency +
                  amount +
                  tof +
                  Mgroup +
                  Fgroup +
                  qty3mo +
                  price3mo +
                  qty6mo +
                  price6mo +
                  qty1mo +
                  price1mo +
                  qty12mo +
                  price12mo +
                  qty +
                  orderinrecent3mo : orderinrecent6mo : orderinrecent12mo +
                  orderinrecent3mo : orderinrecent6mo + 
                  onetothreeflagmo +
                  orderinrecent1mo +
                  orderinrecent3mo +
                  orderinrecent6mo +
                  orderinrecent12mo +
                  startedinrecent1mo +
                  startedinrecent3mo +
                  startedinrecent6mo +
                  startedinrecent12mo +
                  PerSales3mo +
                  PerSales6mo +
                  amtuponorders +
                  qtyuponorders +
                  amtuponqty +
                  ordersupontof +
                  amtupontof +
                  qtyupontof +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20  +
                  newUserFlag
                ,
                
               data = bookModTrain,
               family = binomial)


#summary(bookglm1)

#Running best subset

bookglm <- step(bookglm1, trace = 0)

summary(bookglm)
```

# Tof = 0 handling

```{r}
summary(bookModwo4)

booktof0 <- bookModwo4[(!is.na(bookModwo4$logtargamt)),]

booktof01 <- booktof0[booktof0$tof == 0,]
summary(booktof01)
dim(booktof01)

predlogtargtof0 = mean(booktof01$logtargamt[booktof01$logtargamt > 0])

predprobtof0 = length(booktof01$logtargamt[booktof01$logtargamt > 0]) / length(booktof01$logtargamt)

```

----------------------------------------------------------------------------------------------------------------------------

# Plotting the data

```{r}

# Plotting the data for the fields which will be considered for linear regression
plot(bookModTrain2[,c('logtargamt','recency','frequency')])

plot(bookModTrain2[,c('logtargamt','tof', 'amount' )])

```


# Linear Regression begins here
```{r}

dim(bookModTrain2)

bookModTrain2 <- bookModTrain2[bookModTrain2$tof > 0,]

booklm1 <- lm(logtargamt ~ 
                 recency +
                  frequency +
                  amount +
                  tof +
                  Mgroup +
                  Fgroup +
                  qty3mo +
                  price3mo +
                  qty6mo +
                  price6mo +
                  qty1mo +
                  price1mo +
                  qty12mo +
                  price12mo +
                  qty +
                  orderinrecent3mo : orderinrecent6mo : orderinrecent12mo +
                  orderinrecent3mo : orderinrecent6mo + 
                  onetothreeflagmo +
                  orderinrecent1mo +
                  orderinrecent3mo +
                  orderinrecent6mo +
                  orderinrecent12mo +
                  startedinrecent1mo +
                  startedinrecent3mo +
                  startedinrecent6mo +
                  startedinrecent12mo +
                  PerSales3mo +
                  PerSales6mo +
                  amtuponorders +
                  qtyuponorders +
                  amtuponqty +
                  ordersupontof +
                  amtupontof +
                  qtyupontof +

                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20 +
                  
                  newUserFlag
                ,
                
               data = bookModTrain2)

summary(booklm1)

booklm <- step(booklm1, trace = 0)

summary(booklm)

```

# Model Diagnostics

```{r}

plot(booklm) 
plot(booklm, c(4,6))
str(booklm)
plot(cooks.distance(booklm))
plot(hat(model.matrix(booklm)))

```

# Reruning the regression removing outliers and influential observations

```{r}

4/ (280-14)
bookModTrain3 <- bookModTrain2

# 4 data points have been rmoved on the basis of cook's distance
bookModTrain3 <- bookModTrain3[(cooks.distance(booklm) <= 0.1),]

dim(bookModTrain3)

booklm1r <- lm(logtargamt ~ 
                 recency +
                  frequency +
                  amount +
                  tof +
                  Mgroup +
                  Fgroup +
                  qty3mo +
                  price3mo +
                  qty6mo +
                  price6mo +
                  qty1mo +
                  price1mo +
                  qty12mo +
                  price12mo +
                  qty +
                  orderinrecent3mo : orderinrecent6mo : orderinrecent12mo +
                  orderinrecent3mo : orderinrecent6mo + 
                  onetothreeflagmo +
                  orderinrecent1mo +
                  orderinrecent3mo +
                  orderinrecent6mo +
                  orderinrecent12mo +
                  startedinrecent1mo +
                  startedinrecent3mo +
                  startedinrecent6mo +
                  startedinrecent12mo +
                  PerSales3mo +
                  PerSales6mo +
                  amtuponorders +
                  qtyuponorders +
                  amtuponqty +
                  ordersupontof +
                  amtupontof +
                  qtyupontof +

                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20 +
                  
                  newUserFlag
                ,
                
               data = bookModTrain3)

summary(booklm1r)

booklmr <- step(booklm1r, trace = 0)

summary(booklmr)



```

# Model diagnostics and reruning the model

```{r}

# Checking diagnostics

plot(booklmr) 
plot(booklmr, c(4,6))
str(booklmr)
plot(cooks.distance(booklmr))
plot(hat(model.matrix(booklmr)))

# Checking VIFs
library(faraway)

bookModTrain3$int1 = bookModTrain3$Mhistory19 * bookModTrain3$Fhistory19

vif(bookModTrain3[,c('frequency','amount','qty3mo','price3mo','amtuponorders','amtuponqty','ordersupontof','amtupontof','qtyupontof','Mhistory19','Fconthist20', 'Mconthist20', 'int1')])

str(bookModTrain3)

booklm1r2 <- lm(logtargamt ~ 
                  frequency +
                amount +
                qty3mo +
                price3mo +
                amtuponorders +
                amtuponqty +
                #ordersupontof +
                #amtupontof +
                #qtyupontof +
                Mhistory19 +
                #Fconthist20 + 
                Mhistory19:Fhistory19,
                  
                  
                  data = bookModTrain3
                  )


summary(booklm1r2)

vif(bookModTrain3[,c('frequency','amount','qty3mo','price3mo','amtuponorders','Mhistory19', 'int1','amtuponqty')])


```



# Testing on test data

```{r}

# Logistic model: bookglm

# Linear regression : booklm1r2

predProb = predict(bookglm, newdata=)
adjProb = exp(predProb) / (1 + exp(predProb))

booktest <- booktest[,c('id', 'logtargamt')]
#str(booktesttest)

booktesttest <- bookMod[(is.na(bookMod$logtargamt)),]
booktesttest$logtargamt <- NULL

booktesttest <- merge(booktesttest, booktest, by = 'id', all.x = TRUE, all.y = FALSE)

#summary(booktesttest)

booktesttest$successFlag[booktesttest$logtargamt > 0] <- 1

#table(factor(booktesttest$successFlag))

bookpredicttestmain <- predict(bookglm, newdata = booktesttest, type = 'response')

# Calculating CCRs for various p star values

p <- seq(0, 1, 0.001)

CCR <- p
j <- 0

for (i in p)
{
  tab1 = table(factor(booktesttest$successFlag, levels = c(0,1)),
              factor(bookpredicttestmain > i, levels = c(FALSE,TRUE)))
  #print(tab1)
  
  CCR[j] = ((tab1[1,1] / (tab1[1,1] + tab1[1,2])) + (tab1[2,2] / (tab1[2,1] + tab1[2,2])))/2
  
  j = j + 1
}


plot(p,CCR, xlim = c(0,1))

# Looking at the ROC curve

plot.roc(booktesttest$successFlag, bookpredicttestmain, xlab = "Specificity")


```


## Model Evaluation Criteria

```{r}


# 1. Statistical Criterion

booktest <- booktest[,c('id', 'logtargamt')]
str(booktest) # 25,402 responses

booktesttest <- bookMod[(is.na(bookMod$logtargamt)),]


booktesttest$logtargamt <- NULL

booktesttest <- merge(booktesttest, booktest, by = 'id', all.x = TRUE, all.y = FALSE)

summary(booktesttest)

bookpredicttestmainLogR <- predict(bookglm, newdata = booktesttest, type = 'response')

length(bookpredicttestmainLogR)

summary(booktesttest)

bookpredicttestmainLinR <- predict(booklm, newdata = booktesttest)

bookpredicttestmainLinR[bookpredicttestmainLinR > 10]

abc <- exp(bookpredicttestmainLinR) - 1

EvaluationTable <- booktesttest[,c('id','logtargamt')]

EvaluationTable$logtargamtLogResp <- bookpredicttestmainLogR

EvaluationTable$logtargamtLinResp <- bookpredicttestmainLinR

dim(EvaluationTable)
str(EvaluationTable)
summary(EvaluationTable)

EvaluationTable$ActualTarg <- exp(EvaluationTable$logtargamt) - 1

EvaluationTable$RespTarg <- exp(EvaluationTable$logtargamtLinResp) - 1

EvaluationTable$PredRespTarg <- EvaluationTable$RespTarg * EvaluationTable$logtargamtLogResp

SSEP <- sum((EvaluationTable$PredRespTarg - EvaluationTable$ActualTarg)^2)

paste("Sum of squared errors of prediction: ", SSEP)


# Financial Criterion

summary(EvaluationTable)
top500EvalTable <- head(EvaluationTable[order(EvaluationTable$PredRespTarg, decreasing = TRUE),],500)
TotalPayoff <- sum(top500EvalTable$ActualTarg)

paste("Total payoff based on the top 500 customers based on the predicted data: $", TotalPayoff)
paste("Total payoff as a % of total: ", (TotalPayoff / sum(EvaluationTable$ActualTarg) * 100.0))

i = seq(1,dim(EvaluationTable)[1])

payoffTable <- EvaluationTable[,c(1,2,3)] 
names(payoffTable) <- c('NoOfElements','Payoff', 'STP')

for(i in seq(1,dim(EvaluationTable)[1]))
{
  topEvalTable <- head(EvaluationTable[order(EvaluationTable$PredRespTarg, decreasing = TRUE),],i)
  payoffTable$NoOfElements[i] <- i
  payoffTable$Payoff[i] <- sum(topEvalTable$ActualTarg)
  payoffTable$STP[i] <- payoffTable$Payoff[i] * 0.25 - i 
}

plot(payoffTable$NoOfElements, payoffTable$STP)
payoffTable$NoOfElements[payoffTable$STP == max(payoffTable$STP)]

paste("Maximum number of customers to reach out to maximize the short term profit:", payoffTable$NoOfElements[payoffTable$STP == max(payoffTable$STP)])
paste("Maximum short term profit: $", max(payoffTable$STP))
```

