---
title: "MSiA401Team1Project"
author: "Team 1"
date: "11 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading the files and importing libraries

```{r}
rm(list = ls())
library(pROC)
library(lubridate)
book <- read.csv("book.csv")
ordersall <- read.csv("ordersall.csv", stringsAsFactors = FALSE)
booktest <- read.csv("booktest.csv", stringsAsFactors = FALSE)
summary(book)
summary(ordersall)
```

# Coding success flag in the book database and dropping unnecessary variables
```{r}
book$successFlag <- 0
book$successFlag[book$logtargamt > 0] = 1

book$Mgroup <- book$Mfiction1 + 
  book$Mcartoons5 + 
  book$Mart12 + 
  book$Mhistory19 + 
  book$Mtravelguides31 + 
  book$Mhobby40 + 
  book$Mconthist20 + 
  book$Mnature41

book$Fgroup <- book$Ffiction1 + 
  book$Fcartoons5 + 
  book$Fart12 + 
  book$Fhistory19 + 
  book$Ftravelguides31 + 
  book$Fhobby40 + 
  book$Fconthist20 + 
  book$Fnature41
```
# Extracting features from the ordersall dataset
```{r}
#Converting the orderdate column to date format
typeof(ordersall$orddate)
head(ordersall)
ordersall$orddate <- dmy(ordersall$orddate)
head(ordersall)

#Creating a total quantity database
ordersalltot <- ordersall
aggordersalltot <- ordersalltot[,c('id', 'qty')]
aggordersalltot <- aggregate(aggordersalltot, by = list(aggordersalltot$id), FUN = sum)
aggordersalltot$id <- aggordersalltot$Group.1
aggordersalltot <- aggordersalltot[,c('id', 'qty')]
summary(aggordersalltot)

#Creating a 1 months database
ordersall1mo <- ordersall[ordersall$orddate >= dmy('1-Jul-14'),]

#summary(ordersall1mo)
aggordersall1mo <- ordersall1mo[,c('id', 'qty', 'price')]
aggordersall1mo <- aggregate(aggordersall1mo, by = list(aggordersall1mo$id), FUN = sum)
aggordersall1mo$id <- aggordersall1mo$Group.1
aggordersall1mo <- aggordersall1mo[,c('id', 'qty', 'price')]
summary(aggordersall1mo)
names(aggordersall1mo) <- c('id', 'qty1mo', 'price1mo')

#Creating a 3 months database
ordersall3mo <- ordersall[ordersall$orddate >= dmy('1-May-14'),]
aggordersall3mo <- ordersall3mo[,c('id', 'qty', 'price')]
aggordersall3mo <- aggregate(aggordersall3mo, by = list(aggordersall3mo$id), FUN = sum)
aggordersall3mo$id <- aggordersall3mo$Group.1
aggordersall3mo <- aggordersall3mo[,c('id', 'qty', 'price')]
summary(aggordersall3mo)

#Creating a 6 months database
ordersall6mo <- ordersall[ordersall$orddate >= dmy('1-Feb-14'),]
aggordersall6mo <- ordersall6mo[,c('id', 'qty', 'price')]
aggordersall6mo <- aggregate(aggordersall6mo, by = list(aggordersall6mo$id), FUN = sum)
aggordersall6mo$id <- aggordersall6mo$Group.1
aggordersall6mo <- aggordersall6mo[,c('id', 'qty', 'price')]
summary(aggordersall6mo)

#Creating a 12 months database
ordersall12mo <- ordersall[ordersall$orddate >= dmy('1-Aug-13'),]
aggordersall12mo <- ordersall12mo[,c('id', 'qty', 'price')]
aggordersall12mo <- aggregate(aggordersall12mo, by = list(aggordersall12mo$id), FUN = sum)
aggordersall12mo$id <- aggordersall12mo$Group.1
aggordersall12mo <- aggordersall12mo[,c('id', 'qty', 'price')]
summary(aggordersall12mo)
names(aggordersall12mo) <- c('id', 'qty12mo', 'price12mo')
```

#Merging and creating new features

```{r}
bookMod <- merge(book, aggordersall3mo, by = 'id', all.x = TRUE, all.y = FALSE)
bookMod <- merge(bookMod, aggordersall6mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('3mo','6mo'))
bookMod <- merge(bookMod, aggordersall1mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('','1mo'))
bookMod <- merge(bookMod, aggordersall12mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('','12mo'))
bookMod <- merge(bookMod, aggordersalltot, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('','tot'))


bookMod$qty3mo[is.na(bookMod$qty3mo)] <- 0
bookMod$qty6mo[is.na(bookMod$qty6mo)] <- 0
bookMod$qty1mo[is.na(bookMod$qty1mo)] <- 0
bookMod$qty12mo[is.na(bookMod$qty12mo)] <- 0
bookMod$price3mo[is.na(bookMod$price3mo)] <- 0
bookMod$price6mo[is.na(bookMod$price6mo)] <- 0
bookMod$price1mo[is.na(bookMod$price1mo)] <- 0
bookMod$price12mo[is.na(bookMod$price12mo)] <- 0
bookMod$qty[is.na(bookMod$qty)] <- 0

#Creating a new user flag
bookMod$newUserFlag <- 0
bookMod$newUserFlag[bookMod$tof == 0] <- 1

#Creating a onetothreemoflag month flag
bookMod$onetothreeflagmo <- 0  
bookMod$onetothreeflagmo[((bookMod$qty3mo - bookMod$qty1mo) > 0) & (bookMod$qty1mo == 0)] <- 1

#Creating a flag to track if the user has ordered in the past 1 month
bookMod$orderinrecent1mo <- 0
bookMod$orderinrecent1mo[(bookMod$recency <= 30) && (bookMod$tof > 0)] <- 1

#Creating a flag to track if the user has ordered in the past 3 months
bookMod$orderinrecent3mo <- 0
bookMod$orderinrecent3mo[(bookMod$recency <= 90) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user has ordered in the past 6 months
bookMod$orderinrecent6mo <- 0
bookMod$orderinrecent6mo[(bookMod$recency <= 180) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user has ordered in the past 12 months
bookMod$orderinrecent12mo <- 0
bookMod$orderinrecent12mo[(bookMod$recency <= 360) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user started using the website in the past 30 days
bookMod$startedinrecent1mo <- 0
bookMod$startedinrecent1mo[(bookMod$tof <= 30) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user started using the website in the past 90 days
bookMod$startedinrecent3mo <- 0
bookMod$startedinrecent3mo[(bookMod$tof <= 90) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user started using the website in the past 180 days
bookMod$startedinrecent6mo <- 0
bookMod$startedinrecent6mo[(bookMod$tof <= 180) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user started using the website in the past 360 days
bookMod$startedinrecent12mo <- 0
bookMod$startedinrecent12mo[(bookMod$tof <= 360) & (bookMod$tof != 0)] <- 1

#Creating a field to calculate % of total sales in the past 3 month
bookMod$PerSales3mo <- bookMod$price3mo / bookMod$amount * 100.0
bookMod$PerSales3mo[is.na(bookMod$PerSales3mo)] <- 0

#Creating a field to calculate % of total sales in the past 6 month
bookMod$PerSales6mo <- bookMod$price6mo / bookMod$amount * 100.0
bookMod$PerSales6mo[is.na(bookMod$PerSales6mo)] <- 0

bookMod$amtuponorders <- bookMod$amount / bookMod$frequency
bookMod$qtyuponorders <- bookMod$qty / bookMod$frequency
bookMod$amtuponqty <- bookMod$amount / bookMod$qty
bookMod$ordersupontof <- bookMod$frequency / bookMod$tof
bookMod$amtupontof <- bookMod$amount / bookMod$tof
bookMod$qtyupontof <- bookMod$qty / bookMod$tof

bookMod$amtuponorders[is.na(bookMod$amtuponorders)] <- 0
bookMod$qtyuponorders[is.na(bookMod$qtyuponorders)] <- 0
bookMod$amtuponqty[is.na(bookMod$amtuponqty)] <- 0
bookMod$ordersupontof[is.na(bookMod$ordersupontof)] <- 0
bookMod$amtupontof[is.na(bookMod$amtupontof)] <- 0
bookMod$qtyupontof[is.na(bookMod$qtyupontof)] <- 0
```
# Removing Outliers

```{r}
bookModwo <- bookMod
bookModwo2 <- bookModwo[bookModwo$amtuponorders < 1000,]
bookModwo3 <- bookModwo2[bookModwo2$qtyuponorders < 40,]
bookModwo4 <- bookModwo3[bookModwo3$amtuponqty < 600,]
plot(bookModwo4$amtuponorders)
plot(bookModwo4$qtyuponorders)
plot(bookModwo4$amtuponqty)

perRemoval = (dim(bookMod)[1] - dim(bookModwo4)[1]) / dim(bookMod)[1] * 100
```

#Preliminary logistic regression model on the created dataset

```{r}
# Separating out the test from training dataset
bookModTrain <- bookModwo4[!is.na(bookModwo4$logtargamt),]


# Removing tof = 0
bookModTrain <- bookModTrain[bookModTrain$tof > 0,]

str(bookModTrain)

InitialFactor = dim(bookModTrain[bookModTrain$successFlag == 1,])[1]


#---------------------------------------------------------------------------------
#This data will be used for linear regression
bookModTrain2 <- subset(bookModTrain, logtargamt > 0)
dim(bookModTrain2)
#---------------------------------------------------------------------------------

set.seed(100)
# Bootstraping the data
w = ifelse(bookModTrain$successFlag == 0, 0.8 / length(which(bookModTrain$successFlag == 0)) * 10000, 0.2 / length(which(bookModTrain$successFlag == 1)) * 10000)
bs_idx = sample(seq_len(nrow(bookModTrain)), nrow(bookModTrain), prob=w, replace = TRUE)
bookModTrain = bookModTrain[bs_idx, ]

ScalingFactor <- 1.0 * (dim(bookModTrain[bookModTrain$successFlag == 1,])[1] - InitialFactor) / InitialFactor

# Running the logistic regression
bookglm1 <- glm(as.factor(successFlag) ~ 
                  recency +
                  frequency +
                  amount +
                  tof +
                  Mgroup +
                  Fgroup +
                  qty3mo +
                  price3mo +
                  qty6mo +
                  price6mo +
                  qty1mo +
                  price1mo +
                  qty12mo +
                  price12mo +
                  qty +
                  orderinrecent3mo : orderinrecent6mo : orderinrecent12mo +
                  orderinrecent3mo : orderinrecent6mo + 
                  onetothreeflagmo +
                  orderinrecent1mo +
                  orderinrecent3mo +
                  orderinrecent6mo +
                  orderinrecent12mo +
                  startedinrecent1mo +
                  startedinrecent3mo +
                  startedinrecent6mo +
                  startedinrecent12mo +
                  PerSales3mo +
                  PerSales6mo +
                  amtuponorders +
                  qtyuponorders +
                  amtuponqty +
                  ordersupontof +
                  amtupontof +
                  qtyupontof +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20
                ,
                
               data = bookModTrain,
               family = binomial)


#Running best subset
bookglm <- step(bookglm1, trace = 0)


# Checking VIFs
library(faraway)
bookModTrain$int1 = bookModTrain$Mhistory19*bookModTrain$Fhistory19
bookModTrain$int2 = bookModTrain$Mhealth35*bookModTrain$Fhealth35
bookModTrain$int3 = bookModTrain$Fconthist20*bookModTrain$Mconthist20
vif(bookModTrain[,c('recency','frequency','tof','Mgroup','Fgroup','qty6mo','price6mo','qty1mo','price12mo','qty','orderinrecent3mo','orderinrecent6mo','orderinrecent12mo','startedinrecent3mo','startedinrecent6mo','startedinrecent12mo','PerSales3mo','PerSales6mo','amtuponqty','qtyupontof','Mhistory19','Mcartoons5','Mconthist20','Fconthist20','int1','int2','int3')])


#Refitting the model considering the VIFs into consideration and remiving the insignificant variables from the earlier model
bookglm <- glm(as.factor(successFlag) ~ 
                 recency +
                  frequency +
                  tof +
                  Mgroup +
                  Fgroup +
                  qty6mo +
                  price6mo +
                  qty1mo +
                  price12mo +
                  qty +
                  orderinrecent3mo +
                  orderinrecent6mo +
                  orderinrecent12mo +
                  startedinrecent3mo +
                  startedinrecent6mo +
                  startedinrecent12mo +
                  PerSales3mo +
                  PerSales6mo +
                  amtuponqty +
                  qtyupontof +
                  Mhistory19 +
                  Mcartoons5 +
                  Mhistory19:Fhistory19 +
                  Fhealth35:Mhealth35 +
                  Mconthist20:Fconthist20

                ,
                
               data = bookModTrain,
               family = binomial)

summary(bookglm)


#New VIFs
vif(bookModTrain[,c('recency','frequency','tof','Mgroup','Fgroup','qty6mo','price6mo','qty1mo','price12mo','qty','orderinrecent3mo','orderinrecent6mo','orderinrecent12mo','startedinrecent3mo','startedinrecent6mo','startedinrecent12mo','PerSales3mo','PerSales6mo','amtuponqty','qtyupontof','Mhistory19','Mcartoons5','int1','int2','int3')])


#Model Diagnostics
bookModTrainTest <- bookModwo4[!is.na(bookModwo4$logtargamt),]
bookModTrainTest <- bookModTrainTest[bookModTrainTest$tof > 0,]

predProbTrain <- predict(bookglm, bookModTrainTest, type = "response")

# Plotting ROC Curve

plot.roc(bookModTrainTest$successFlag, predProbTrain, xlab = "Specificity")
auc(roc(bookModTrainTest$successFlag, predProbTrain))

```

# Plotting the data

```{r}
# Plotting the data for the fields which will be considered for linear regression
plot(bookModTrain2[,c('logtargamt','recency','frequency','tof', 'amount')])

```


# Linear Regression begins here
```{r}

# Removing observations with tof > 0
bookModTrain2 <- bookModTrain2[bookModTrain2$tof > 0,]

booklm1 <- lm(logtargamt ~ 
                 recency +
                  frequency +
                  amount +
                  tof +
                  Mgroup +
                  Fgroup +
                  qty3mo +
                  price3mo +
                  qty6mo +
                  price6mo +
                  qty1mo +
                  price1mo +
                  qty12mo +
                  price12mo +
                  qty +
                  orderinrecent3mo : orderinrecent6mo : orderinrecent12mo +
                  orderinrecent3mo : orderinrecent6mo + 
                  onetothreeflagmo +
                  orderinrecent1mo +
                  orderinrecent3mo +
                  orderinrecent6mo +
                  orderinrecent12mo +
                  startedinrecent1mo +
                  startedinrecent3mo +
                  startedinrecent6mo +
                  startedinrecent12mo +
                  PerSales3mo +
                  PerSales6mo +
                  amtuponorders +
                  qtyuponorders +
                  amtuponqty +
                  ordersupontof +
                  amtupontof +
                  qtyupontof +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20
                ,
                
               data = bookModTrain2)

#Running best Subset
booklm <- step(booklm1, trace = 0)
summary(booklm)
```

# Model Diagnostics

```{r}
plot(booklm) 
plot(booklm, c(4,6))
str(booklm)
plot(cooks.distance(booklm))
plot(hat(model.matrix(booklm)))
```

# Reruning the regression removing outliers and influential observations

```{r}
bookModTrain3 <- bookModTrain2

# Removing data points with cook's distance > 0.1
bookModTrain3 <- bookModTrain3[(cooks.distance(booklm) <= 0.1),]
dim(bookModTrain3)

# Checking VIFs
library(faraway)
bookModTrain3$int1 = bookModTrain3$Mhistory19 * bookModTrain3$Fhistory19
vif(bookModTrain3[,c('frequency','amount','qty3mo','price3mo','amtuponorders','amtuponqty','ordersupontof','amtupontof','qtyupontof','Mhistory19','Fconthist20', 'Mconthist20', 'int1')])

# As we cans see, we have a few variables with VIFs which are significantly high

#Rerunning the model removing the outliers based on cook's distance, considering the VIFs into consideration and removing the insignificant variables from the earlier model
booklm1r2 <- lm(logtargamt ~ 
                  frequency +
                amount +
                amtuponorders +
                amtuponqty +
                Mhistory19 +
                Mhistory19:Fhistory19,
                  
                  data = bookModTrain3
                  )

summary(booklm1r2)

```

# Model diagnostics and reruning the model

```{r}
# Checking diagnostics
plot(booklm1r2) 
plot(booklm1r2, c(4,6))

# Checking VIFs
library(faraway)
bookModTrain3$int1 = bookModTrain3$Mhistory19 * bookModTrain3$Fhistory19
vif(bookModTrain3[,c('frequency','amount','amtuponorders','Mhistory19', 'int1','amtuponqty')])
```



## Tof = 0 handling

```{r}

booktof0 <- bookModwo4[(!is.na(bookModwo4$logtargamt)),]
booktof01 <- booktof0[booktof0$tof == 0,]

dim(booktof01)

#the predicted value for non-responders
predlogtargtof0 = mean(booktof01$logtargamt[booktof01$logtargamt > 0])

#the probability for non-responders
predprobtof0 = length(booktof01$logtargamt[booktof01$logtargamt > 0]) / length(booktof01$logtargamt)
```

## Prediction of Responder and Non-responders and Final results
```{r}
booktesttest <- bookMod[(is.na(bookMod$logtargamt)),]
dim(booktesttest)

booktest <- booktest[,c('id', 'logtargamt')]
str(booktesttest)

booktesttest$logtargamt <- NULL
booktesttest <- merge(booktesttest, booktest, by = 'id', all.x = TRUE, all.y = FALSE)
booktesttest$successFlag[booktesttest$logtargamt > 0] <- 1

#the final probability for each responder
booktest1 = booktesttest[booktesttest$tof > 0,]
booktest0 = booktesttest[booktesttest$tof == 0,]

dim(booktest1)
dim(booktest0)

predProb = predict(bookglm, newdata = booktest1, type = 'response')
predScale = -log(ScalingFactor) + log(predProb / (1+predProb))
adjProb = exp(predScale) / (1 + exp(predScale)) #the final probability for each responder
length(adjProb)

booktest1$successFlag <- 0
booktest1$successFlag[booktest1$logtargamt > 0] <- 1
length(booktest1$successFlag)
length(adjProb)

plot.roc(booktest1$successFlag, adjProb, xlab = "Specificity")
auc(roc(booktest1$successFlag, adjProb))



#the predicted value for responders
predictedvals = predict(booklm1r2, newdata=booktest1)
predictednewvals = rep(predlogtargtof0,length(booktest0$id))

#the final predictions for responders
result1 = booktest1[c('id', 'logtargamt')]
result1$predlog = adjProb * predictedvals

#final predictions for non-responders
result0 = booktest0[c('id', 'logtargamt')]
result0$predlog = predprobtof0 * predictednewvals

#combine the two results 
result = rbind(result0, result1)
result$predtarget = exp(result$predlog) - 1
result$target = exp(result$logtargamt) - 1

#adjust for extreme values 
#change any predicted value larger than 2 times the max amount in the training set (3 data points) to 2 times the max amount of the training set 
result[result$predtarget > 2*max(exp(bookModTrain3$logtargamt)-1),]$predtarget = max(exp(bookModTrain3$logtargamt)-1) 


# 1. Statistical Criterion

#sse
result_ssep = sum((result$predtarget - result$target)^2)
result_ssep
rmse = sqrt(result_ssep / nrow(result)-1)

# 2. Financial Criterion

#payoff
payoff = sum(result[order(result$predtarget, decreasing = TRUE),]$target[c(1:500)]) / sum(result[order(result$target, decreasing = TRUE),]$target[c(1:500)])
payoff

#Optimal customer count to reach out to

payoffTable <- result[,c(1,2,3)] 
names(payoffTable) <- c('NoOfElements','Payoff', 'STP')

for(i in seq(1,dim(result)[1]))
{
  topEvalTable <- head(result[order(result$predtarget, decreasing = TRUE),],i)
  payoffTable$NoOfElements[i] <- i
  payoffTable$Payoff[i] <- sum(topEvalTable$target)
  payoffTable$STP[i] <- payoffTable$Payoff[i] * 0.25 - i 
}

plot(payoffTable$NoOfElements, payoffTable$STP)

plot(payoffTable$NoOfElements, payoffTable$STP, xlim = c(0,3000), ylim = c(0,1500))
payoffTable$NoOfElements[payoffTable$STP == max(payoffTable$STP)]
paste("Maximum number of customers to reach out to maximize the short term profit:", payoffTable$NoOfElements[payoffTable$STP == max(payoffTable$STP)])
paste("Maximum short term profit: $", max(payoffTable$STP))

#payoff for top # of customers per the above calculations
payoff2 = sum(result[order(result$predtarget, decreasing = TRUE),]$target[c(1:payoffTable$NoOfElements[payoffTable$STP == max(payoffTable$STP)])]) / sum(result[order(result$target, decreasing = TRUE),]$target[c(1:payoffTable$NoOfElements[payoffTable$STP == max(payoffTable$STP)])])
payoff2


```


