---
title: "401 project"
author: "Yiwei Zhang"
date: "11/12/2018"
output: html_document
---

```{r}
#helper function to get the accuracy of prediction and actual
accy = function(actual, predicted) {
  accy = sum(actual == predicted)/length(actual)
  return(accy)
}
rmse = function(actual, predicted){
  res = sqrt(sum((actual - predicted)^2) / length(actual))
  return(res)
}

library(pROC)
library(lubridate)
```

```{r}
#read in the data
book = read.csv("~/Desktop/MSiA401/project/book.csv")
book_train = read.csv("~/Desktop/MSiA401/project/booktrain.csv")
book_test = read.csv("~/Desktop/MSiA401/project/booktest.csv")
ordersall = read.csv("~/Desktop/MSiA401/project/ordersall.csv", stringsAsFactors = FALSE)
```

-------------------------------------------------
Category Features
-------------------------------------------------

```{r}
#combination of similar categories
book$Mgroup <- book$Mfiction1 + 
  book$Mcartoons5 + 
  book$Mart12 + 
  book$Mhistory19 + 
  book$Mtravelguides31 + 
  book$Mhobby40 + 
  book$Mconthist20 + 
  book$Mnature41

book$Fgroup <- book$Ffiction1 + 
  book$Fcartoons5 + 
  book$Fart12 + 
  book$Fhistory19 + 
  book$Ftravelguides31 + 
  book$Fhobby40 + 
  book$Fconthist20 + 
  book$Fnature41

book$successFlag <- 0
book$successFlag[book$logtargamt > 0] = 1
```

-------------------------------------------------
Order Features
-------------------------------------------------
Create dataframe for 3 month, 6 month, 1 year, 3 year
```{r}
#order data
ordersall$orddate <- dmy(ordersall$orddate)

#Creating a total quantity database
ordersalltot <- ordersall
aggordersalltot <- ordersalltot[,c('id', 'qty')]
aggordersalltot = aggregate(aggordersalltot$qty, by = list(aggordersalltot$id), FUN = "sum")
aggordersalltot$id <- aggordersalltot$Group.1
aggordersalltot$qty = aggordersalltot$x
aggordersalltot <- aggordersalltot[,c('id', 'qty')]
summary(aggordersalltot)

#Creating a 3 months database
ordersall3mo <- ordersall[ordersall$orddate >= dmy('1-May-14'),]
#summary(ordersall3mo)
aggordersall3mo <- ordersall3mo[,c('id', 'qty', 'price')]
aggordersall3mo <- aggregate(aggordersall3mo[,-1], by = list(aggordersall3mo$id), FUN = "sum")
aggordersall3mo$id <- aggordersall3mo$Group.1
aggordersall3mo <- aggordersall3mo[,c('id', 'qty', 'price')]
summary(aggordersall3mo)

#Creating a 6 months database
ordersall6mo <- ordersall[ordersall$orddate >= dmy('1-Feb-14'),]
#summary(ordersall6mo)
aggordersall6mo <- ordersall6mo[,c('id', 'qty', 'price')]
aggordersall6mo <- aggregate(aggordersall6mo[,-1], by = list(aggordersall6mo$id), FUN = "sum")
aggordersall6mo$id <- aggordersall6mo$Group.1
aggordersall6mo <- aggordersall6mo[,c('id', 'qty', 'price')]
summary(aggordersall6mo)

#Creating a 1 year database
ordersall1yr <- ordersall[ordersall$orddate >= dmy('1-Aug-13'),]
#summary(ordersall6mo)
aggordersall1yr <- ordersall1yr[,c('id', 'qty', 'price')]
aggordersall1yr <- aggregate(aggordersall1yr[,-1], by = list(aggordersall1yr$id), FUN = "sum")
aggordersall1yr$id <- aggordersall1yr$Group.1
aggordersall1yr <- aggordersall1yr[,c('id', 'qty', 'price')]
summary(aggordersall1yr)

#Creating a 2 year database
ordersall2yr <- ordersall[ordersall$orddate >= dmy('1-Aug-12'),]
#summary(ordersall6mo)
aggordersall2yr <- ordersall2yr[,c('id', 'qty', 'price')]
aggordersall2yr <- aggregate(aggordersall2yr[,-1], by = list(aggordersall2yr$id), FUN = "sum")
aggordersall2yr$id <- aggordersall2yr$Group.1
aggordersall2yr <- aggordersall2yr[,c('id', 'qty', 'price')]
summary(aggordersall2yr)

#Creating a 3 year database
ordersall3yr <- ordersall[ordersall$orddate >= dmy('1-Feb-11'),]
#summary(ordersall6mo)
aggordersall3yr <- ordersall3yr[,c('id', 'qty', 'price')]
aggordersall3yr <- aggregate(aggordersall3yr[,-1], by = list(aggordersall3yr$id), FUN = "sum")
aggordersall3yr$id <- aggordersall3yr$Group.1
aggordersall3yr <- aggordersall3yr[,c('id', 'qty', 'price')]
summary(aggordersall3yr)
```

Merge to get order variables
```{r}
#merge the data and do the order features
bookMod <- merge(book, aggordersall3mo, by = 'id', all.x = TRUE, all.y = FALSE)
bookMod <- merge(bookMod, aggordersall6mo, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('3mo', '6mo'))
bookMod <- merge(bookMod, aggordersall1yr, by = 'id', all.x = TRUE, all.y = FALSE)
bookMod <- merge(bookMod, aggordersall2yr, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('1yr','2yr'))
bookMod <- merge(bookMod, aggordersall3yr, by = 'id', all.x = TRUE, all.y = FALSE)
bookMod <- merge(bookMod, aggordersalltot, by = 'id', all.x = TRUE, all.y = FALSE, suffixes = c('3yr','tot'))
names(bookMod)[names(bookMod) == "price"] = "price3yr"
names(bookMod)[names(bookMod) == "qtytot"] = "qty"

bookMod$qty3mo[is.na(bookMod$qty3mo)] <- 0
bookMod$qty6mo[is.na(bookMod$qty6mo)] <- 0
bookMod$qty1yr[is.na(bookMod$qty1yr)] <- 0
bookMod$qty3yr[is.na(bookMod$qty3yr)] <- 0
bookMod$qty2yr[is.na(bookMod$qty2yr)] <- 0
bookMod$price3mo[is.na(bookMod$price3mo)] <- 0
bookMod$price6mo[is.na(bookMod$price6mo)] <- 0
bookMod$price1yr[is.na(bookMod$price1yr)] <- 0
bookMod$price2yr[is.na(bookMod$price2yr)] <- 0
bookMod$price3yr[is.na(bookMod$price3yr)] <- 0
bookMod$qty[is.na(bookMod$qty)] <- 0

#Creating a new user flag
bookMod$newUserFlag <- 0
bookMod$newUserFlag[bookMod$tof == 0] <- 1

#Creating a flag to track if the user has ordered in the past 1 month
bookMod$orderinrecent1mo <- 0
bookMod$orderinrecent1mo[(bookMod$recency <= 30) && (bookMod$tof > 0)] <- 1

#Creating a flag to track if the user has ordered in the past 3 months
bookMod$orderinrecent3mo <- 0
bookMod$orderinrecent3mo[(bookMod$recency <= 90) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user has ordered in the past 6 months
bookMod$orderinrecent6mo <- 0
bookMod$orderinrecent6mo[(bookMod$recency <= 180) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user has ordered in the past 1 year
bookMod$orderinrecent1yr <- 0
bookMod$orderinrecent1yr[(bookMod$recency <= 365) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user started using the website in the past 30 days
bookMod$startedinrecent1mo <- 0
bookMod$startedinrecent1mo[(bookMod$tof <= 30) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user started using the website in the past 90 days
bookMod$startedinrecent3mo <- 0
bookMod$startedinrecent3mo[(bookMod$tof <= 90) & (bookMod$tof != 0)] <- 1

#Creating a flag to track if the user started using the website in the past 180 days
bookMod$startedinrecent6mo <- 0
bookMod$startedinrecent6mo[(bookMod$tof <= 180) & (bookMod$tof != 0)] <- 1

#Creating a field to calculate % of total sales in the past 3 month
bookMod$PerSales3mo <- bookMod$price3mo / bookMod$amount * 100.0
bookMod$PerSales3mo[is.na(bookMod$PerSales3mo)] <- 0

#Creating a field to calculate % of total sales in the past 6 month
bookMod$PerSales6mo <- bookMod$price6mo / bookMod$amount * 100.0
bookMod$PerSales6mo[is.na(bookMod$PerSales6mo)] <- 0

#Creating a field to calculate % of total sales in the past 1 year
bookMod$PerSales1yr <- bookMod$price1yr / bookMod$amount * 100.0
bookMod$PerSales1yr[is.na(bookMod$PerSales1yr)] <- 0

#trend of ordering price in the past 6 month, 1, 2 and 3 years; not using 3 month because most price3mo and qty3mo are 0s
bookMod$price_change_6mo = ifelse(bookMod$price1yr - bookMod$price6mo < 0.1 | (2*bookMod$price6mo - bookMod$price1yr) / (bookMod$price1yr - bookMod$price6mo) == Inf, 2*bookMod$price6mo - bookMod$price1yr, (2*bookMod$price6mo - bookMod$price1yr) / (bookMod$price1yr - bookMod$price6mo))

bookMod$price_change_1yr = ifelse(bookMod$price2yr - bookMod$price1yr < 0.1, 2*bookMod$price1yr - bookMod$price2yr, (2*bookMod$price1yr - bookMod$price2yr) / (bookMod$price2yr - bookMod$price1yr))

bookMod$price_change_2yr = ifelse(bookMod$price3yr - bookMod$price2yr < 0.1, 2*bookMod$price2yr - bookMod$price3yr, (2*bookMod$price2yr - bookMod$price3yr) / (bookMod$price3yr - bookMod$price2yr))

#trend of ordering quantity in the past 6 month, 1, 2 and 3 years; not using 3 month because most price3mo and qty3mo are 0s
bookMod$qty_change_6mo = ifelse(bookMod$qty1yr - bookMod$qty6mo < 0.1, 2*bookMod$qty6mo - bookMod$qty1yr, (2*bookMod$qty6mo - bookMod$qty1yr) / (bookMod$qty1yr - bookMod$qty6mo))
bookMod$qty_change_1yr = ifelse(bookMod$qty2yr - bookMod$qty1yr < 0.1, 2*bookMod$qty1yr - bookMod$qty2yr, (2*bookMod$qty1yr - bookMod$qty2yr) / (bookMod$qty2yr - bookMod$qty1yr))
bookMod$qty_change_2yr = ifelse(bookMod$qty3yr - bookMod$qty3yr < 0.1, 2*bookMod$qty2yr - bookMod$qty3yr, (2*bookMod$qty2yr - bookMod$qty3yr) / (bookMod$qty3yr - bookMod$qty3yr))

#amount, quantity per order
bookMod$amtuponorders <- bookMod$amount / bookMod$frequency
bookMod$qtyuponorders <- bookMod$qty / bookMod$frequency
#amount per quantity
bookMod$amtuponqty <- bookMod$amount / bookMod$qty

bookMod$amtuponorders[is.na(bookMod$amtuponorders)] <- 0
bookMod$qtyuponorders[is.na(bookMod$qtyuponorders)] <- 0
bookMod$amtuponqty[is.na(bookMod$amtuponqty)] <- 0

summary(bookMod)
```

----------------------------------------------
Category features
----------------------------------------------

test and train split
```{r}
#train test split based on given data
btrain = merge(book, book_train, by = "id")
btrain = subset(btrain, select = -c(logtargamt.y,X) )
btest = merge(book, book_test, by = "id")
btest = subset(btest, select = -c(logtargamt.x, X))
```

In general category variables are of pretty importance 
```{r}
#create binary variables for classification model
btrain$buy = ifelse(btrain$logtargamt.x > 0, 1, 0)
btest$buy = ifelse(btest$logtargamt.y > 0, 1, 0)

#random forest for variable importance 
library(caret)
ctrl = trainControl(method="cv", number = 5) 
class.rf = train(as.factor(buy) ~ ., data = subset(btrain, select = -c(id, logtargamt.x)), method = "rf", trControl = ctrl, tuneGrid = expand.grid(mtry = c(5, 10, 20, 30, 50)))
rf.accy = accy(btest$buy, predict(class.rf, newdata=btest))
rf.accy #0.9618
```

```{r}
#accuracy and variable importance 
#summary(class.rf)
varImp(class.rf)
```

If only uses category variables, added M44, 12, 36 and F50 to the list. 
```{r}
#only use categories variables
btrain_cat = btrain[,c(1, 2, 7:67)]
class.rf2 = train(as.factor(buy) ~ ., data = subset(btrain_cat, select = -c(id, logtargamt.x)), method = "rf", trControl = ctrl, tuneGrid = expand.grid(mtry = c(5, 10, 20, 30, 50)))
rf.accy2 = accy(btest$buy, predict(class.rf2, newdata=btest))
```

So, adding categories do not in extra valuable information ....
```{r}
#accuracy and feature importance
rf.accy2 #0.9606, same as null model (classifying everything to 0); sample ratio is 0.9606
varImp(class.rf2)
```

```{r}
#if only use non-categories
class.rf0 = train(as.factor(buy) ~ ., data = subset(btrain, select = c(recency, frequency, amount, tof, buy)), method = "rf", trControl = ctrl, tuneGrid = expand.grid(mtry = c(1,2,3,4)))
rf.accy0 = accy(btest$buy, predict(class.rf0, newdata=btest)) #0.9625, same as full model
rf.accy0
```

```{r}
#regression, use only non-0 data
ctrl = trainControl(method="cv", number = 5) 
bregre = btrain[which(btrain$logtargamt.x > 0),]
bregre_test = btest[which(btest$logtargamt.y > 0),]
regre.rf = train(logtargamt.x ~ ., data = subset(bregre, select = -c(id, buy)), method = "rf", trControl = ctrl, tuneGrid = expand.grid(mtry = c(5, 10, 20, 30, 50)), importance = T)
rf.rmse.regre = rmse(bregre_test$logtargamt.y, predict(regre.rf, newdata=bregre_test))
rf.rmse.regre #0.70
```

```{r}
varImp(regre.rf)
```

```{r}
#only uses non-categories
regre.rf0 = train(logtargamt.x ~ ., data = subset(bregre, select = c(recency, frequency, amount, tof, logtargamt.x)), method = "rf", trControl = ctrl, tuneGrid = expand.grid(mtry = c(1,2,3,4)), importance = T)
rf.regre0 = rmse(bregre_test$logtargamt.y, predict(regre.rf0, newdata=bregre_test)) 
rf.regre0 #0.6987
```
```{r}
varImp(regre.rf0)
```
```{r}
#interaction among the important categories
#20 14 35 19 8 5
#10 is really low so deleted
regre.rf2 = train(logtargamt.x ~ recency + frequency + amount + tof + Fconthist20 + Mconthist20 + Fmusic14 + Mmusic14 + Fhealth35  + Mhealth35 + Freligion8 + Mreligion8 + Fhistory19 + Mhistory19 + Mcartoons5 + Fcartoons5 + Fconthist20*Mconthist20 + Fmusic14*Mmusic14 + Fhealth35*Mhealth35 + Freligion8*Mreligion8 + Fcartoons5*Mcartoons5 + Fhistory19*Mhistory19, data = subset(bregre, select = -c(id, buy)), method = "rf", trControl = ctrl, tuneGrid = expand.grid(mtry = c(5, 10, 15, 20)), importance = T)
rf.rmse.regre2 = rmse(bregre_test$logtargamt.y, predict(regre.rf2, newdata=bregre_test))
rf.rmse.regre2 #0.7045
```

Final important categories features
```{r}
varImp(regre.rf2)
```

----------------------------------------------
Modeling
----------------------------------------------
```{r}
#outlier removal

```


```{r}
# Separating out the test from training dataset
#Without outlier removed select this code line
bookModTrain <- bookMod[(!is.na(bookMod$logtargamt)),]
bookModTrain2 <- subset(bookModTrain, logtargamt > 0)

# For replicating the minor dataset
bookModTrainpos <- bookModTrain[bookModTrain$successFlag == 1,]
for (i in seq(1,10))
{
  bookModTrain <- rbind(bookModTrain, bookModTrainpos)
}

# Running the logistic regression
bookglm1 <- glm(successFlag ~ 
                 recency + 
                 frequency + 
                 amount +
                 tof +
                 qty3mo + 
                 price3mo + 
                 qty6mo +
                 price6mo +
                 orderinrecent1mo +
                 orderinrecent3mo +
                 orderinrecent6mo +
                 startedinrecent1mo +
                 startedinrecent3mo +
                 startedinrecent6mo +
                 PerSales3mo +
                 PerSales6mo +
                 Mgroup +
                 Fgroup +
                 amtuponorders +
                 qtyuponorders +
                 amtuponqty +
                  price_change_1yr +
                  price_change_2yr +
                  price_change_6mo +
                  qty_change_2yr + 
                  qty_change_1yr + 
                  qty_change_6mo +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20
                ,
               data = bookModTrain,
               family = binomial)
summary(bookglm1)

#Running best subset
bookglm <- step(bookglm1)
summary(bookglm)
```


--------------------------------------
Linear Regression
--------------------------------------
```{r}
booklm1 <- lm(logtargamt ~ 
                 recency + 
                 frequency + 
                 amount +
                 tof +
                 qty3mo + 
                 price3mo + 
                 qty6mo +
                 price6mo +
                 orderinrecent1mo +
                 orderinrecent3mo +
                 orderinrecent6mo +
                 startedinrecent1mo +
                 startedinrecent3mo +
                 startedinrecent6mo +
                 PerSales3mo +
                 PerSales6mo +
                 Mgroup +
                 Fgroup +
                 amtuponorders +
                 qtyuponorders +
                 amtuponqty +
                  price_change_1yr +
                  price_change_2yr +
                  price_change_6mo +
                  qty_change_2yr + 
                  qty_change_1yr + 
                  qty_change_6mo +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20
                ,
               data = bookModTrain2)
summary(booklm1)
booklm <- step(booklm1, trace=0)
summary(booklm)
```

After remove outliers
```{r}
bookModTrain3 <- bookModTrain2
bookModTrain3 <- bookModTrain3[(cooks.distance(booklm) <= 0.025) & (hat(model.matrix(booklm)) <= 0.25),]

booklm1r <- lm(logtargamt ~ 
                 recency + 
                 frequency + 
                 amount +
                 tof +
                 qty3mo + 
                 price3mo + 
                 qty6mo +
                 price6mo +
                 orderinrecent1mo +
                 orderinrecent3mo +
                 orderinrecent6mo +
                 startedinrecent1mo +
                 startedinrecent3mo +
                 startedinrecent6mo +
                 PerSales3mo +
                 PerSales6mo +
                 Mgroup +
                 Fgroup +
                 amtuponorders +
                 qtyuponorders +
                 amtuponqty +
                  price_change_1yr +
                  price_change_2yr +
                  price_change_6mo +
                  qty_change_2yr + 
                  qty_change_1yr + 
                  qty_change_6mo +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20
                ,
               data = bookModTrain3)
summary(booklm1r)
booklm2 <- step(booklm1r, trace=0)
summary(booklm2)
```

Standarized Data
```{r}
bookModTrain_tof = bookModTrain2[bookModTrain2$tof > 0,]
book_scale = data.frame(scale(bookModTrain_to, center=TRUE, scale = TRUE))
booklm_scale <- lm(logtargamt ~ 
                 recency + 
                 frequency + 
                 amount +
                 tof +
                 qty3mo + 
                 price3mo + 
                 qty6mo +
                 price6mo +
                 orderinrecent3mo +
                 orderinrecent6mo +
                 startedinrecent1mo +
                 startedinrecent3mo +
                 startedinrecent6mo +
                 PerSales3mo +
                 PerSales6mo +
                 Mgroup +
                 Fgroup +
                 amtuponorders +
                 qtyuponorders +
                 amtuponqty +
                  price_change_1yr +
                  price_change_2yr +
                  price_change_6mo +
                  qty_change_2yr + 
                  qty_change_1yr + 
                  qty_change_6mo +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20
                ,
               data = book_scale)
summary(booklm_scale)
booklm_scale <- step(booklm_scale, trace=0)
summary(booklm_scale)
```

Outliers removal for scaled data
```{r}
#remove outliers for scaled data
book_scale2 <- book_scale
book_scale2 <- book_scale2[(cooks.distance(booklm_scale) <= 0.012) & (hat(model.matrix(booklm_scale)) <= 0.079),]

booklm_scale2 <- lm(logtargamt ~ 
                 recency + 
                 frequency + 
                 amount +
                 tof +
                 qty3mo + 
                 price3mo + 
                 qty6mo +
                 price6mo +
                 orderinrecent3mo +
                 orderinrecent6mo +
                 startedinrecent1mo +
                 startedinrecent3mo +
                 startedinrecent6mo +
                 PerSales3mo +
                 PerSales6mo +
                 Mgroup +
                 Fgroup +
                 amtuponorders +
                 qtyuponorders +
                 amtuponqty +
                  price_change_1yr +
                  price_change_2yr +
                  price_change_6mo +
                  qty_change_2yr + 
                  qty_change_1yr + 
                  qty_change_6mo +
                  Mhistory19 +
                  Mcartoons5 +
                  Mconthist20 +
                  Fhistory19:Mhistory19 +
                  Fhealth35:Mhealth35 +
                  Fconthist20 +
                  Mcartoons5:Fcartoons5 +
                  Mconthist20:Fconthist20
                ,
               data = book_scale2)
summary(booklm_scale2)
booklm_scale2 <- step(booklm_scale2, trace=0)
summary(booklm_scale2)
```




